% !TeX spellcheck = en_US
% !TeX encoding = UTF-8
% !TeX root = ../document.tex

\chapter{Introduction}


\section{Units and Abbreviations}
Throughout this work, we will use the natural unit system. This means that the speed of light and the Planck constant are fixed to $1$:
\begin{equation*}
    \hbar = c = 1
\end{equation*}
It follows that the only unit needed to express most other physical quantities is the unit of energy. Our choice here is electron volts. Too keep numbers in a reasonable range, we mostly use gigaelectronvolts: $\SI{1}{\giga\eV} = \SI{1.60218e-10}{\joule}$.
Two exceptions to this are the cross section and luminosity, which are expressed in femtobarn and inverse femtobarn respectively: $\SI{1}{\femto\barn} = \SI{e-43}{\meter\squared}$.

\section{Particle Physics}
Finding the building blocks of matter has interested mankind for thousands of years. Ancient Greek philosophers already contemplated about the basic building blocks of matter. They coined the term \emph{atom} for what they thought to be indivisible. Since then, our understanding of the smallest parts and what holds them together has vastly improved. This field of science is nowadays called \emph{Elementary Particle Physics}. As of 2017, the state of particle physics research is summed up in the so called \emph{Standard Model of Particle Physics}.

In the following sections, we will give a short introduction to the Standard Model, its theoretical foundations and open questions. Motivated by these questions, we will introduce possible extensions to the Standard Model in the next chapter.

This chapter is based on \cite{Hebbeker:SkriptzurElementarteilchenphysik} if not explicitly stated otherwise.

\section{The Standard Model}
The Standard Model is a description of matter and its interactions, except gravity. It is based on the theoretical foundations of quantum field theory.


\subsection{Particles and Interactions}
According to the Standard Model, elementary particles are point-like indivisible structureless objects, much like the atoms were to the ancient Greeks. These objects are characterized by physical observables such as mass, charge and spin.

\begin{figure}
    \centering
    \tikzset{font=\small,
        edge from parent fork down,
        level distance=3em,
        sibling distance=0.3em,
        every node/.style=
        {
            align=center,
        },
        every leaf node/.style={
            circle,
%            rectangle,rounded corners,
            draw,           
            very thick,
            inner sep=0,
            minimum size=1.4em,
            text height=0.5em,
            text depth=0,
        }
    }
    \begin{tikzpicture}
    \Tree
    [.{Elementary Particles}
        [.Fermions
            [.Leptons
                [.{charged}
                    {\Pe}
                    {\Pmu}
                    {\Ptau}
                ]
                [.{neutrinos}
                    {\Pnue}
                    {\Pnum}
                    {\Pnut}
                ]
            ]
            [.Quarks
%               [.{1st gen.}
%                   {\Pup}
%                   {\Pdown}
%               ]
%               [.{2nd gen.}
%                   {\Pcharm}
%                   {\Pstrange}
%               ]
%               [.{3rd gen.}
%                   {\Pbottom}
%                   {\Ptop}
%               ]
                [.{up-type}
                    {\Pup}
                    {\Pcharm}
                    {\Pbottom}
                ]
                [.{down-type}
                    {\Pdown}
                    {\Pstrange}
                    {\Ptop}
                ]
            ]
        ]
        [.Bosons
            {\Pgamma}
            {\Pgluon}
            [.{charged}
                {\PW}
                {\PZ}
            ]
            {\PHiggs}
        ]
    ]
    \end{tikzpicture}
    \caption{Nested groups of elementary particles. The actual particle symbols are indicated by circles.}
    \label{fig:particle_groups}
\end{figure}

Figure \ref{fig:particle_groups} shows how elementary particles can be separated into several groups. 
At first level there are two groups: \emph{fermions} and \emph{bosons}.

Fermions are elementary particles with a spin of $\nicefrac{1}{2}$. They follow Fermi-Dirac statistics and have to obey Pauli's exclusion principle which states that no more than one particle can occupy a certain state characterized by its quantum numbers.
The group of fermions can be further divided into quarks and leptons.
There are six quarks: up (\Pup), down (\Pdown), charm (\Pcharm), strange (\Pstrange), bottom (\Pbottom) and top (\Ptop). Quarks carry color charge and are the only objects to take part in the strong interaction.
In addition, there are three lepton families: \emph{electron} (\Pe), \emph{muon} (\Pmu) and \emph{tau} (\Ptau). Each charged lepton has a very light, electrically neutral \emph{neutrino} associated with it.

The second group of particles are called bosons. They mediate interactions (also called \emph{forces}) and posses an integer spin. Bosons follow Bose-Einstein statistics and may thus occupy the same quantum state.
Each kind of boson is responsible for one kind of elementary force: the electrodynamic force is mediated by the \emph{photon}, the strong force by the \emph{gluon} and the weak force by the \emph{\PZ} and \emph{\PW} bosons.
In addition, there is the \emph{Higgs boson} (\PH), which uses the Higgs mechanism to give mass to all mentioned massive elementary particles.\footnote{Note that this does not apply to composite particles (like protons) which gain mask mostly through binding energy.}


\begin{table}
    \centering
    \begin{tabular}{l r s[table-unit-alignment = left]}
        \toprule
        Lepton & \multicolumn{2}{l}{Mass} \\
        \midrule
        \Pe & 511 & \keV \\
        \Pnue & < 2 & \eV \\
        \Pmu & 106 & \MeV \\
        \Pnum & < 2 & \eV \\
        \Ptau & 1.78 & \GeV \\
        \Pnut & < 2 & \eV \\
        \bottomrule
    \end{tabular}
    \begin{tabular}{l r s[table-unit-alignment = left]}
        \toprule
        Quark & \multicolumn{2}{l}{Mass} \\
        \midrule
        \Pup & 2 & \MeV \\
        \Pdown & 5 & \MeV \\
        \Pstrange & 96 & \MeV \\
        \Pcharm & 1.3 & \GeV \\
        \Pbottom & 4.18 & \GeV \\
        \Ptop & 173 & \GeV \\
        \bottomrule
    \end{tabular}
    \begin{tabular}{l r s[table-unit-alignment = left]}
        \toprule
        Boson & \multicolumn{2}{l}{Mass} \\
        \midrule
        \Pgamma & 0 &  \\
        \Pgluon & 0 &  \\
        \PW & 80.4 & \GeV \\
        \PZ & 91.2 & \GeV \\
        \PH & 125 & \GeV \\
        \bottomrule
    \end{tabular}
    \caption{Known elementary particles and their masses\cite{ParticleDataGroup:ReviewParticlePhysics}. Note that for layouting purposes, quarks and leptons have been put side-by-side, even though there is no known connection between the number of lepton and quark families.}
\end{table}

In terms of conventional matter, one can conclude that fermions make up matter and bosons bind matter together.

%\begin{table}
%   \begin{tabular}{ l l l l l }
%       Interaction & Couples to & Mediator & Theory & Strength \\
%       \hline
%       Strong & Color charge & Gluon (\Pgluon) & Chromodyn. & \num{1} \\
%       Electromagnetic & Electric charge & Photon (\Pphoton) & Electrodyn. & \num{e-3} \\
%       Weak & Flavor charge & \PW and \PZ & Flavordyn. & \num{e-14} \\
%       Gravitational & & & & \num{e-43} \\
%   \end{tabular}
%   \caption{Overview over elementary interactions. The Standard Model does not describe gravity, it is only included in the table for strength comparison.\cite{Griffiths:IntroductiontoElementaryParticles}}
%\end{table}
%
\subsection{QFD}
Weak force

change flavor

lightest particles are stable

\subsection{Open Questions}
Since the discovery of the Higgs boson in 2012, all particles predicted by the Standard Model have been experimentally observed. The Standard Model has not been disproven and multiple Nobel prices have been awarded.
One could assume that our understanding of nature is now complete, but this is not the case:
In the recent decades, several observations have been made that demand  theories beyond the Standard Model. In the following sections, we want to discuss a few of them.

%Observations: gravitational waves (2016), neutrino masses

\subsubsection{Astrophysical Observations}
One astrophysical research method is to probe whether observed gravitational effects can be completely accounted for by visible matter distributions. For this purpose, the rotation curves of galaxies have been measured and shown to be incompatible with simulations of visible matter only. Also, fluctuations in gravitational lensing without a visible origin has been observed.\cite{Bertone:Particledarkmatter,Peebles:Cosmologicalconstantdark}
Furthermore, several experiments have probed the cosmic microwave background for anisotropies. These data have subsequently been compared to the matter distributions obtained by simulations of the early universe according to the $\Uplambda$CDM model. The results indicated that only about \SI{5}{\percent} of the universe is made up of conventional (baryonic) matter.\cite{Planck:Planck2015results}
% https://arxiv.org/pdf/1003.0939v2.pdf

All of these findings suggest that there must be a large amount of matter in the universe that only interacts gravitationally. Because of its electromagnetic invisibility, it is called \emph{Dark Matter}.
Since the Standard Model does not provide a fitting Dark Matter candidate, this is an active field of research also at particle collider experiments.

\subsubsection{Neutrino Masses}
So far, no right-handed neutrinos have been detected. Thus the Standard Model was crafted in a way that neutrinos only occur as left-handed particles.
Recent observations of neutrino oscillations\cite{KamLAND:ReactorAntineutrinoMeasurement,DoubleChooz:Improvedmeasurementsneutrino,IceCube:Determiningneutrinooscillation,DayaBay:NewMeasurementAntineutrino} indicate that neutrinos may be massive. For the theory to remain renormalizable, neutrinos must thus also exist in their right-handed state\cite{Klinkhamer:NeutrinomassStandard}, which demands for an extension or modification of the Standard Model.

\subsubsection{Theoretical Considerations}
From the theoretical point of view, the open questions are mostly due to incomplete theories: 
First, there is no consensus about unification of the gauge interactions. At certain very high energy scales, the gauge interactions are expected to behave similarly. However, the observed Standard Model parameters do not allow for a common point of convergence\cite{Amaldi:Comparisongrandunified}.
Secondly, the framework of quantum field theory does not allow for a non-renormalizable theory such as quantum gravity. Thus, the question about how to combine particle physics and gravity (or rather general relativity) is still open.\footnote{In the context of collider experiments, the effects of gravity are so weak that they can be neglected.}

Another theoretical motivation is the \emph{mass hierarchy problem}: One of the seemingly indisputable energy scales in nature is the Planck scale at $M_\text{Pl} \approx \SI{e18}{\GeV}$. At this scale, gravity becomes as strong as the gauge interactions and nature is expected to behave differently. However, all processes that have been observed so far appear at the electroweak scale $m_\text{EW} \approx \SI{e3}{\GeV}$. This discrepancy between these two scales challenges the generality of the Standard Model.

% Spin 3/2 particles?

\section{Extensions of the Standard Model}


\section{Experiments}
Since Rutherford's experiments in the early 20th century, there have been an abundance of particle physics experiments using the approach of particle collisions. Today, most scattering experiments can be classified in two groups: Fixed target experiments and collider experiments. In the former case, accelerated particles are collided with a fixed target, such as Rutherford's gold foils. In the latter case, two beams of particles are made to collide at an interaction point. This allows for a higher center of mass energy and is thus favorable whenever the technical circumstances allow.

In both cases, outgoing scattered particles are registered. Because the Standard Model makes precise predictions about the rate of which certain scattering scenarios occur, counting outgoing particles with certain properties is a very efficient way of testing the theory.

\section{The Large Hadron Collider}
The to date largest particle collider experiment is located underground just below Geneva in Switzerland. It belongs to the European Organization for Nuclear Research CERN\footnote{The acronym CERN stems from research organization's original French name: \textbf{C}onseil \textbf{E}uropéen pour la \textbf{R}echerche \textbf{N}ucléaire} and is called \emph{Large Hadron Collider} (LHC). The LHC is a proton-proton collider with a total circumference of \SI{26.7}{\km}. At two of its interaction points, detectors have been constructed that serve as general purpose experiments: The \emph{Compact Muon Solenoid} (CMS) and the ATLAS experiment\cite{Evans:LHCMachine}. 
This work will focus on the former experiment, the CMS detector.

\section{The Compact Muon Solenoid}



\subsection{Detector Design}
\subsection{Subsystems}
\subsection{Triggering}
\subsection{The Computing Grid}